---
title: "MachineLearningModels"
format: html
---

```{r}
library(baseballr)
library(tidyverse)
library(tidymodels)
library(tibble)
library(dplyr)
library(lubridate)
library(olsrr)
library(readr)
library(rsample)
library(tidymodels)
library(glmnet)
library(rpart)
library(rpart.plot)
library(caret)
library(caTools) 
library(randomForest) 
library(ggplot2)
library(ISLR2)
library(pls)
library(class)
library(ROCR)
```

## Summary Write-Up

-   Timeline is on track at the moment, since I was planning on over half my ML models done by the middle of this month, I have gotten most of the "big" ones put together and have begun analysis relating to them.

    -   I plan on comparing the MSE scores, and looking at the weights for all of my models as well. What is becoming significant?

    -   Multicollinearity is a major issue (will show a example of why below)

# Frequentist and ML Approaches for Baseball

## Establish Dataset

```{r}
#Load Data
data_final = read_csv('data_smaller.csv')
set.seed(1)
```

```{r}
#Sending file

write.csv(data_complete_na, "data_completed_na.csv") 
```

```{r}
head(data_final)
```

```{r}
data_final %>% count(pitch_type)
```

## Removing Redundant Pitches

```{r}
#So we're gonna drop some data from the pitch_type column, FC is "fielders choice" which really isn't a pitch. it also had some limited connections to the data.

#I am also dropping pitches that had low amounts of existence to begin with. (Need to find research to back this)
#drop_list <- ("CS","FA","EP",'FC','FO','KN')

data_final <- data_final %>% filter(., !pitch_type %in% c("CS",'FA','EP','FC','FO','KN','PO','SC','SV'))
```

## Multiple Linear Regression

```{r}
#Using "+" for each term, "*" used to mark interaction terms.
#Adding in release_pos_z since that is a well-used statistics for ball release position due to its ability to track the abll movement towards the catcher
lm.fit <- lm(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type + release_pos_z , data = data_final)

summary(lm.fit)
```

```{r}
lm.fit <- lm(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + release_pos_z + pitch_type, data = data_complete_na)

summary(lm.fit)
```

### Checking Multiple Linear Regression Assumptions?

```{r}
#Checking model assumptions post-creation
#What is happening LOL
plot(lm.fit)
```

## Weighted Multiple Linear Regression

Coming soon...

## Ridge Regression

```{r}
#Had a issue with NA values, so removing them here before moving on.
data_complete_na <- data_final %>%
  select(surgery, release_speed, arm_angle, spin_axis, 
         release_spin_rate, pitch_type, release_pos_z) %>%
  na.omit()
```

```{r}
#We will be using the glmnet package for this section, following instructions from the ISLP textbook.
#We will begin with ridge regression

#Set up X and Y
x <- model.matrix(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type + release_pos_z, data = data_complete_na)[,-1]
y <- data_complete_na$surgery
#Set up grid and fitting the ridge regression model
grid <- 10^seq(10, -2, length = 100)
ridge.mod <- glmnet(x,y, alpha = 0, lambda = grid)

#Size of our matrix
#dim(coef(ridge.mod))
```

### Best MSE for Ridge Regression

```{r}
#Create our training/testing data
train <- sample(1:nrow(x), nrow(x) * 0.8)  
test <- (-train)
y.test <- y[test]
```

```{r}
ridge.mod <- glmnet(x[train,], y[train], alpha = 0, lambda = grid, thresh = 1e-12)
ridge.pred <- predict(ridge.mod, s = 4, newx = x[test,])
#Test MSE
mean((ridge.pred - y.test)^2)
```

```{r}
#Yea this was made?
par(mar = c(5, 4, 4, 8))
plot(ridge.mod)

last_coef <- coef(ridge.mod)[, ncol(coef(ridge.mod))]
last_coef <- last_coef[-1]

axis(4, at = last_coef, labels = names(last_coef), las = 1, cex.axis = 0.7)
```

```{r}
#Find the best lambda
cv.out <- cv.glmnet(x[train,], y[train], alpha = 0)
plot(cv.out)
```

```{r}
bestlam <- cv.out$lambda.min
bestlam
```

```{r}
#MSE associated with the best lambda
ridge.pred <- predict(ridge.mod, s = bestlam, newx = x[test,])
mean((ridge.pred - y.test)^2)
```

```{r}
#Refitting our ridge regression on the full dataset using the lambda chosen for use. (The best lambda we found)
#Weights of the values
out <- glmnet(x, y, alpha = 0)
predict(out, type = "coefficients", s = bestlam)[1:13,]
```

### Multiple Linear Regression (Ridge)

```{r}
data_filtered <- data_final[data_final$pitch_type %in% c("FS","FF", "KC", "SL"), ]

lm.tuned.ridge.fit <- lm(surgery ~ arm_angle + release_spin_rate + pitch_type, data = data_complete_na)

summary(lm.tuned.ridge.fit)
```

```{r}
plot(lm.tuned.ridge.fit)
```

### Ridge Regression Confusion Matrix

This model is even worse when compared to the PCR Confusion Matrix, and struggles to balance the Type I and Type II error well to receive good results.

It does a very poor job at predicting when a pitcher will not have surgery

```{r}
#Get the prediction class for our model and set as a factor.
ridge.pred.prob <- predict(ridge.mod, s = 4, newx = x[test,], type = "response")

# Convert probabilities to class predictions
# Make sure you know what your surgery levels are
data_complete_na$surgery <- as.factor(data_complete_na$surgery)

ridge_pred_class <- ifelse(ridge.pred.prob > 0.285, 
                          levels(data_complete_na$surgery)[2], 
                          levels(data_complete_na$surgery)[1])
ridge_pred_class <- as.factor(ridge_pred_class)

# Get actual test values
actual_class <- data_complete_na$surgery[test]

# Create confusion matrix
ridge_conf <- confusionMatrix(ridge_pred_class, 
                             actual_class,
                             positive = levels(actual_class)[2])

print(ridge_conf)

```

```{r}
#ROC curve 
pred <- prediction(ridge.pred.prob, actual_class)
perf <- performance(pred, "sens", "spec")
plot(perf,
     avg = "threshold",
     colorize = TRUE,
     lwd = 3,
     main = "Sensitivity/Specificity Plot for Ridge Regression"
    )
```

## Lasso Regression

```{r}
#Fitting a lasso model
lasso.mod <- glmnet(x[train,], y[train],alpha = 1, lambda = grid, importance = TRUE)
plot(lasso.mod, label = TRUE)
```

```{r}
par(mar = c(5, 4, 4, 12))
plot(lasso.mod, xlim = c(2, 5), ylim = c(-0.15, 0.05))

last_coef <- coef(lasso.mod)[, ncol(coef(lasso.mod))]
last_coef <- last_coef[-1]

# --- spread labels that are too close together ---
spread_labels <- function(positions, min_gap = 0.008) {
  pos <- positions
  sorted_idx <- order(pos)
  pos_sorted <- pos[sorted_idx]
  
  # Iteratively push labels apart
  for (iter in 1:100) {
    moved <- FALSE
    for (i in 2:length(pos_sorted)) {
      if ((pos_sorted[i] - pos_sorted[i-1]) < min_gap) {
        mid <- (pos_sorted[i] + pos_sorted[i-1]) / 2
        pos_sorted[i-1] <- mid - min_gap / 2
        pos_sorted[i]   <- mid + min_gap / 2
        moved <- TRUE
      }
    }
    if (!moved) break
  }
  
  result <- pos
  result[sorted_idx] <- pos_sorted
  return(result)
}

# Get spread label positions
label_pos <- spread_labels(last_coef, min_gap = 0.012)

# Draw axis with spread labels + segments connecting to true positions
axis(4, at = label_pos, labels = names(last_coef), las = 1, cex.axis = 0.7, tick = FALSE)

# Tick marks at TRUE coefficient positions
axis(4, at = last_coef, labels = FALSE, tcl = -0.3)

# Connecting lines from true position to spread label
segments(
  x0 = par("usr")[2],          # right edge of plot
  y0 = last_coef,              # true value
  x1 = par("usr")[2] * 1.01,  # just outside plot
  y1 = label_pos,              # spread label position
  col = "gray60", lwd = 0.7, xpd = TRUE
)
```

```{r}
#importance_weights(lasso.mod)
```

```{r}
#Perform cross-validation to compute the test error.
cv.out <- cv.glmnet(x[train,], y[train], alpha = 1)
plot(cv.out)
```

```{r}
#Find our MSE
bestlam.lasso <- cv.out$lambda.min
lasso.pred <- predict(lasso.mod, s = bestlam.lasso, newx = x[test,])
mean((lasso.pred - y.test)^2)
```

```{r}
#CV has occured, we get our coefs.
out <- glmnet(x,y,alpha = 1, lambda = grid)
lasso.coef <- predict(out, type = 'coefficients', s = bestlam.lasso)[1:13,]
lasso.coef
```

### Lasso Regression Confusion Matrix

```{r}
#Get the prediction class for our model and set as a factor.
lasso.pred.prob <- predict(lasso.mod, s = 4, newx = x[test,], type = "response")

# Convert probabilities to class predictions
# Make sure you know what your surgery levels are
#data_complete_na$surgery <- as.factor(data_complete_na$surgery)

lasso_pred_class <- ifelse(lasso.pred.prob > 0.285, 
                          levels(data_complete_na$surgery)[2], 
                          levels(data_complete_na$surgery)[1])
lasso_pred_class <- as.factor(ridge_pred_class)

# Get actual test values
actual_class <- data_complete_na$surgery[test]

# Create confusion matrix
lasso_conf <- confusionMatrix(lasso_pred_class, 
                             actual_class,
                             positive = levels(actual_class)[2])

print(lasso_conf)
```

## Principle Component Regression

```{r}
#Model creation
pcr.fit <- pcr(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type + release_pos_z , data = data_complete_na , scale = TRUE , validation = "CV")
```

```{r}
#Model Summary
#Gonna need to get some help understanding these results well.
summary(pcr.fit)
```

```{r}
#Cross-Validation MSE being plotted below
#Due to the small number of compoents used, a better model might do better with less "noise" of data included.
validationplot(pcr.fit, val.type = "MSEP")
```

### Training/Testing PCR

```{r}
#Perform a PCR on the trainign data and evaluate its test set performance
pcr.fit <- pcr(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type , data = data_final, subset = train, scale = TRUE, validation = 'CV')
validationplot(pcr.fit, val.type = "MSEP")
```

```{r}
#Computing to find the test MSE
pcr.pred <- predict(pcr.fit, x[test,], ncomp = 6)
mean((pcr.pred - y.test)^2)
```

```{r}
#Fit the PCR to the full data set, using M = 4
#Ok so, I think I need to have more data explained by 4 components if I really want this to work...
pcr.fit <- pcr(y ~ x, scale = TRUE, ncomp = 6)
summary(pcr.fit)
```

### PCR Confusion Matrix

```{r}
#Setup
#Establish Surgery as a Factor
data_complete_na$surgery <- as.factor(data_complete_na$surgery)
pcr_pred_prob <- predict(pcr.fit, x[test,], ncomp = 6)

#Binary classification
#changing that value to .3 allows for the model to have more freedom in choosing it's threshold value for prediction.
pcr_pred_class <- ifelse(pcr_pred_prob > 0.30, 
                         levels(data_complete_na$surgery)[2], 
                         levels(data_complete_na$surgery)[1])
pcr_pred_class <- as.factor(pcr_pred_class)

#Testing values
actual_class <- data_complete_na$surgery[test]

#Confusion Matrix
pcr_conf <- confusionMatrix(pcr_pred_class, 
                            actual_class, 
                            positive = levels(actual_class)[2])

print(pcr_conf)
#Moving the threshold of the pcr_pred_class allowed me to find a better balance for the confusion matrix.
```

## Partial Least Squares Regression

```{r}
#Using the plsr() libary like the pls() function. Slight different here
pls.fit <- plsr(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type + release_pos_z , data = data_final , subset = train, scale = TRUE, validation = "CV")
summary(pls.fit)
```

```{r}
#Cross-validation error 
validationplot(pls.fit, val.type =  "MSEP")
```

```{r}
#MSE
pls.pred <- predict(pls.fit, x[test,],ncomp = 13)
mean((pls.pred - y.test)^2)
```

```{r}
#Avoid this rn
pls.fit <- plsr(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type , data = data_final , scale = TRUE, ncomp = 12)
summary(pls.fit)
```

### Partial Least Squares Confusion Matrix

This one has the same setup as the confusion matrix for principle component regression.

```{r}
#Establish Surgery as a Factor
#Uncomment to run if needed.
data_complete_na$surgery <- as.factor(data_complete_na$surgery)
pls_pred_prob <- predict(pls.fit, x[test,], ncomp = 13)

#Binary classification
#changing that value to .3 allows for the model to have more freedom in choosing it's threshold value for prediction.
pls_pred_class <- ifelse(pls_pred_prob > 0.30, 
                         levels(data_complete_na$surgery)[2], 
                         levels(data_complete_na$surgery)[1])
pls_pred_class <- as.factor(pcr_pred_class)

#Testing values
actual_class <- data_complete_na$surgery[test]

#Confusion Matrix
pls_conf <- confusionMatrix(pls_pred_class, 
                            actual_class, 
                            positive = levels(actual_class)[2])

print(pls_conf)
```

## KNN Regression

```{r}
#Setup process, training and testing data. Setting direction as well.
dim(data_complete_na)
attach(data_complete_na)

#Standardize the data so all variables are given a mean of 0, SD of 1.
standardized.X <- scale(data_complete_na[,-6])
dim(standardized.X)
```

```{r}
#Set up training and testing data. Using the first 20000 points for the testing data.
test <- 1:20000
train.X <- standardized.X[-test,]
test.X <- standardized.X[test,]
train.Y <- surgery[-test]
test.Y <- surgery[test]
```

There is some issue where with using KNN and it being unable to move itself with in the data, have to look into fixing...

```{r}
#Begin KNN predicatiom with K = 1
knn.pred <- knn(train.X, test.X, train.Y, k = 50)
#Should be greater then 0?
mean(test.Y != knn.pred)
mean(test.Y != 0)
```

```{r}
table(knn.pred, test.Y)
```

## Random Forest

```{r}
data_complete_na <- data_complete_na %>%
  mutate(
    pitch_type = as.factor(pitch_type),    # must be factor for randomForest
    surgery    = as.factor(surgery),       # must be factor for classification
    release_speed     = as.numeric(release_speed),
    arm_angle         = as.numeric(arm_angle),
    spin_axis         = as.numeric(spin_axis),
    release_spin_rate = as.numeric(release_spin_rate)
  )
```

```{r}
str(data_complete_na)
```

```{r}
#This is the actual random forest, ignore all the other crap 

library(caret)
library(randomForest)

# Split from data_complete_na consistently
train_index <- createDataPartition(data_complete_na$surgery, p = 0.8, list = FALSE)
train <- data_complete_na[train_index, ]
test  <- data_complete_na[-train_index, ]

# Train model on train (no need for subset= when using train directly)
rf.surgery <- randomForest(
  surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type,
  data       = train,       # ← use train directly instead of subset =
  mtry       = 3,
  importance = TRUE
)

# Predict on test
predictions <- predict(rf.surgery, newdata = test)

# Confusion matrix
confusionMatrix(predictions, test$surgery)

# Variable importance plot
varImpPlot(rf.surgery, main = "Variable Importance for Surgery Prediction")

```

```{r}
bag.baseball <- randomForest(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type , data = data_complete_na, subset = train, mtry = 12, importance = TRUE)

bag.baseball
```

```{r}
#this portion is giving me trouble trying to find MSE, will come back to and debug more later
yhat.bag <- predict(bag.baseball, newdata = data_complete_na[-train,])
data_complete_na.test <- data_complete_na[-train, "surgery"]
```

```{r}
plot(yhat.bag, data_complete_na.test)
abline(0,1)
```

```{r}
mean((yhat.bag - data_complete_na.test)^2)
```

Getting the warning "invalid mtry : reset to within range", will have to investigate where that is coming from and the reason behind its occurrence.

```{r}
#Growing a random forest
rf.baseball <- randomForest(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type , data = data_complete_na, subset = train, mtry = 6, importance = TRUE)

yhat.rf <- predict(rf.baseball, newdata = data_complete_na[-train,])
mean((yhat.rf - data_complete_na.test)^2)
```

```{r}
importance(rf.baseball)
```

```{r}
varImpPlot(rf.baseball)
```

### Random Forest Confusion Matrix

To do...

## Bayes Lab

```{r}
#Data reset here so I have everything straight as I need it
bayes_data$surgery = as.numeric(data_complete_na$surgery)
#bayes_data$surgery = as.numeric(bayes_data$pitch_type)
#levels(bayes_data$surgery) = c('No Surgery','Surgery')

#bayes_data = data_complete_na
bayes_data
```

```{r}
#Import the packages I need
#I got bros email on SPEEDDIAL
library(BayesFactor)
library(brms)
```

```{r}
#I first tried to use anovaBF, but since my predictors are continuous, I was suggested to use regressionBF() instead
#Ok that didn't work, swapped to lmBF. I will also try brms here too.

#Since lmBF (from my reading/understanding) is just the Bayesian form of linear regression, I will have to do some digging into brms and other methods to possibly find a way to get better results.

bayesball.lm = lmBF(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type , data = bayes_data)
```

```{r}
bayesball.lm
```

```{r}
summary(bayesball.lm)
```

```{r}
#This final suggestion is something that Claude just cooked up, I really don't know what is happening here.

# Extract posterior samples for deeper inspection
posterior_samples <- posterior(bayesball.lm, iterations = 1000)
summary(posterior_samples)

# Plot posterior distributions
plot(posterior_samples)
```

### BRMS Stuff

```{r}
data_complete_na$pitch_FF <- as.integer(data_complete_na$pitch_type == "FF")
data_complete_na$pitch_SL <- as.integer(data_complete_na$pitch_type == "SL")
data_complete_na$pitch_FS <- as.integer(data_complete_na$pitch_type == "FS")
data_complete_na$pitch_KC <- as.integer(data_complete_na$pitch_type == "KC")
```

```{r}
bayes.brms <- brm(surgery ~ release_speed + arm_angle + spin_axis + release_spin_rate + pitch_type + release_pos_z + pitch_FF:pitch_SL + pitch_FF:pitch_FS + pitch_FF:pitch_KC , data = data_complete_na, 
                  family = bernoulli(link = 'logit'),
                  chains = 5)
```

```{r}
prior_summary(bayes.brms)
```

```{r}
summary(bayes.brms)
```

```{r}
#Graphics
plot(bayes.brms)
```

```{r}
# Get all plots as a list
bayes_plots <- plot(bayes.brms, ask = FALSE)  # ask = FALSE stops it prompting between pages

# Save each one separately
for (i in seq_along(bayes_plots)) {
  ggsave(
    filename = paste0("bayes_plot_", i, ".png"),
    plot     = bayes_plots[[i]],
    width    = 14,
    height   = 10,
    dpi      = 300
  )
}
```

```{r}
library(bayesplot)

# Marginal effects of each predictor
#conditional_effects(bayes.brms)

# Specific predictor
conditional_effects(bayes.brms, effects = "release_speed")
conditional_effects(bayes.brms, effects = "pitch_type")
```

## SHAP / XGBoost

```{r}
library(xgboost)
library(iml)

#Dataset we will use here
data_complete_na

#Establish training and test
#train_shap <- as.matrix(data_complete_na[, -1])  # Features
train_shap_labels <- data_complete_na$surgery 
```

```{r}
# Convert to factor first
data_complete_na$pitch_type <- as.factor(data_complete_na$pitch_type)

# Save the mapping with clearer column names
factor_levels <- levels(data_complete_na$pitch_type)
factor_mapping <- data.frame(
  pitch_name    = factor_levels,
  encoded_value = 1:length(factor_levels)
)

print(factor_mapping)

# Then encode and build matrix
data_complete_na$pitch_type <- as.numeric(data_complete_na$pitch_type)
train_shap <- as.matrix(data_complete_na[, -1])

```

```{r}
#Model for xgboost
xgboost.model <- xgboost(x = train_shap, y = train_shap_labels, nrounds = 100, objective = "binary:logistic")
```

```{r}
# Define a custom predict function
predict_function <- function(model, newdata) {
  predict(model, as.matrix(newdata))
}

# Pass it into Predictor$new
predictor.shap <- Predictor$new(
  xgboost.model,
  data = as.data.frame(train_shap),
  y = train_shap_labels,
  predict.function = predict_function
)

# Now try Shapley again
shapley <- Shapley$new(predictor.shap, x.interest = as.data.frame(train_shap[1, , drop = FALSE]))
```

```{r}
shap_data <- shapley$results %>%
  mutate(feature.value = gsub("=.*", "", feature.value)) %>%
  mutate(feature.value = gsub("pitch_type", "Slider (SL)", feature.value))

ggplot(shap_data, aes(x = reorder(feature.value, phi), 
                      y = phi, 
                      fill = ifelse(phi > 0, "Positive", "Negative"))) +
  geom_col() +
  scale_fill_manual(
    values = c("Positive" = "firebrick", "Negative" = "steelblue"),
    name   = "Direction"
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title      = element_text(face = "bold", size = 14),
    axis.text       = element_text(size = 11),
    legend.position = "bottom"
  ) +
  labs(
    title = "SHAP Values for\nSurgery Prediction",
    x     = "Feature",
    y     = "SHAP Value"
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50")
```

## Presentation Supplies

```{r}
data_complete_na %>% count(pitch_type)
```

```{r}
library(ggplot2)
library(dplyr)
data_complete_na %>%
  count(pitch_type) %>%
  mutate(pitch_type = reorder(pitch_type, n)) %>%
  ggplot(aes(x = pitch_type, y = n)) +
  geom_col(fill = "steelblue", width = 0.7) +     
  scale_y_continuous(labels = scales::comma) +
  coord_flip() +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text  = element_text(size = 11),
    panel.grid.major.y = element_blank()
  ) +
  labs(
    title = "Pitch Type Distribution",
    x     = "Pitch Type",
    y     = "Count"
  )
```

```{r}
library(gtsummary)

tbl <- data_complete_na %>%
  tbl_summary(
    by = surgery,                        # your group column (0/1 or "Yes"/"No")
    statistic = list(
      all_continuous()  ~ "{mean} ± {sd}",   # matches the ± format in your table
      all_categorical() ~ "{n} ({p})"         # matches the n (%) format
    ),
    digits = all_continuous() ~ 2,
    missing = "no"
  ) %>%
  add_p() %>%                            # adds the P column
  bold_p(t = 0.05) %>%                   # bolds significant p-values like in your table
  modify_header(label ~ "**Variables**") %>%
  modify_spanning_header(
    c("stat_1", "stat_2") ~ "**Pitching Statistics**"
  )

tbl
```

```{r}
library(gt)
```

```{r}
# Save as HTML (no browser needed)
tbl %>%
  as_gt() %>%
  gt::gtsave("data_table2.html")
```

```{r}
library(gtsummary)

# --- Section 1: Release Metrics ---
tbl1 <- data_complete_na %>%
  select(surgery, release_speed, release_pos_z, release_spin_rate) %>%
  tbl_summary(
    by = surgery,
    statistic = list(all_continuous() ~ "{mean} \u00b1 {sd}"),
    digits = all_continuous() ~ 2,
    missing = "no"
  ) %>%
  add_p(test = list(all_continuous() ~ "kruskal.test")) %>%
  bold_p(t = 0.05)

# --- Section 2: Pitch Mechanics ---
tbl2 <- data_complete_na %>%
  select(surgery, arm_angle, spin_axis, pitch_type) %>%
  tbl_summary(
    by = surgery,
    statistic = list(
      all_continuous()  ~ "{mean} \u00b1 {sd}",
      all_categorical() ~ "{n} ({p})"
    ),
    digits = all_continuous() ~ 2,
    missing = "no"
  ) %>%
  add_p(test = list(
    all_continuous()  ~ "kruskal.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_p(t = 0.05)

# --- Stack and format ---
tbl_final <- tbl_stack(
  tbls = list(tbl1, tbl2),
  group_header = c("Release Metrics", "Pitch Mechanics")
) %>%
  modify_header(label ~ "**Variables**") %>%
  bold_labels()

tbl_final
```

```{r}

library(gtsummary)
library(dplyr)
library(purrr)

# Get all unique pitch types dynamically
pitch_types <- unique(data_complete_na$pitch_type)

# Build one table per pitch type
tbl_list <- map(pitch_types, ~ {
  data_complete_na %>%
    filter(pitch_type == .x) %>%
    select(surgery, release_speed, arm_angle,
           spin_axis, release_spin_rate, release_pos_z) %>%
    tbl_summary(
      by = surgery,
      statistic = list(all_continuous() ~ "{mean} \u00b1 {sd}"),
      digits = all_continuous() ~ 2,
      missing = "no"
    ) %>%
    add_ci(
      method = list(all_continuous() ~ "t.test"),
      style_fun = all_continuous() ~ function(x) style_sigfig(x, digits = 2)
    ) %>%
    add_p(
      test = list(all_continuous() ~ "wilcox.test")
    ) %>%
    bold_p(t = 0.05)
})

# Stack all pitch type tables vertically
tbl_final <- tbl_stack(
  tbls = tbl_list,
  group_header = pitch_types     # uses pitch type name as section divider
) %>%
  modify_header(label ~ "**Variable**") %>%
  bold_labels()

tbl_final
```

```{r}
data_complete_na %>% count(surgery)
```
